On The Benefits of Magnitude Estimation Relevance Assessments for Information Retrieval Evaluation

### 2016-09-14 11_appendix.tex
Attempted to fix the missing answer to RQ2.

Don’t worry about running out of numbers -- there will always be a
larger number than the largest you use, and a smaller number than the
smallest you use.

@@

\begin{itemize}
\item[2.a] First, a statement that expesses a need for information will be
displayed at the top of the screen.  Please read the statement
carefully.  You will be asked a question about the statement, to test
your understanding.

\item[2.b] You will then be asked to rate 8 documents that have been returned
by a search system, in response to the information need statement.
\end{itemize}

@@

>>>>

Don't worry about running out of numbers -- there will always be a
larger number than the largest you use, and a smaller number than the
smallest you use.

@@

\begin{enumerate}
\item[2.a] First, a statement that expresses a need for information will be
displayed at the top of the screen.  Please read the statement
carefully.  You will be asked a question about the statement, to test
your understanding.

\item[2.b] You will then be asked to rate 8 documents that have been returned
by a search system, in response to the information need statement.
\end{enumerate}

@@

% Local Variables:
% TeX-master: "ME-TOIS.tex"
% End:

### 2016-09-14 4_descriptive_stats.tex
another increment...
Attempted to fix the missing answer to RQ2.

Magnitude estimation scores should be approximately
log-normal~\citep{Mar74,moskowitz:1977}. 
%\sm{ks test on the logs does not indicate normality but...}
Figure~\ref{fig:ME-raw-scores} shows the situation for our data, 
plotting the Cumulative Distribution Functions (CDFs) of:
all the scores (top left), all the scores in the $0.001$-$100$ range
(which includes about 99.5\% of the scores), and the scores for each
topic. 
The continuous blue lines are fitted log-normal distributions with the
same mean and standard deviation of the corresponding data.  
From the figure it is clear that the log-normal distribution is a
reasonable approximation of the raw data. 
Differences are generally due to a \emph{round number tendency}
\cite{moskowitz:1977}: in general, people expressing magnitudes using
a numeric scale tend to use some specific values more than others, for
instance integer numbers and powers of ten. 
From the figure it can be seen that our workers tended to use
especially 1, 2, 5, and 10 (the values where ``steps'' can be seen on
the CDF curves --- see the gray vertical dashed lines on the charts),
and also quite a lot of 3, 4, 5, 6 7, 8, 9, 20, 30, 40, 50, 60, 70,
80, 90, and 100. 
The breakdowns on single topics show that topics do have an effect
(for instance, in topics 418 and 421 the scores up to~1 included
account for about 44\% and 23\% of the values respectively), although
the general trend is very similar, as is again demonstrated by the
fitted log-normal distributions.
Topic means ranged from 7.9 to 12.9, and topic standard deviations ranged
from 17.5 to 23.2.

@@

Other normalizations have been proposed, 
for example using the median or the arithmetic mean instead of the
geometric mean (especially when zero is an allowed score, so the
geometric mean cannot be used)~\citet{moskowitz:1977}.
Other approaches include carrying out an additional calibration task
that is aimed at understanding the differences in individual scales;
in our scenario, this might be implemented by exploiting the \nkn and
\hkh documents, or in a more implicit way by relying on the maximal
and minimal scores expressed by each worker.

>>>>

Magnitude estimation scores should be approximately
log-normally distributed~\citep{Mar74,moskowitz:1977}. 
%\sm{ks test on the logs does not indicate normality but...}
Figure~\ref{fig:ME-raw-scores} shows the situation for our data, 
plotting the Cumulative Distribution Functions (CDFs) of:
all the scores (top left), all the scores in the $0.001$-$100$ range
(which includes about 99.5\% of the scores), and the scores for each
topic. 
The continuous blue lines are fitted log-normal distributions with the
same mean and standard deviation of the corresponding data.  
From the figure it is clear that the log-normal distribution is a
reasonable approximation of the raw data. 
Differences are generally due to a \emph{round number tendency}
\cite{moskowitz:1977}: in general, people expressing magnitudes using
a numeric scale tend to use some specific values more than others, for
instance integer numbers and powers of ten. 
From the figure it can be seen that our workers tended to use
especially 1, 2, 5, and 10 (the values where ``steps'' can be seen on
the CDF curves --- see the gray vertical dashed lines on the charts),
and also quite a lot of 3, 4, 5, 6 7, 8, 9, 20, 30, 40, 50, 60, 70,
80, 90, and 100. 
The breakdowns on single topics show that topics do have an effect
(for instance, in topics 418 and 421 the scores up to~1 included
account for about 44\% and 23\% of the values respectively), although
the general trend is very similar, as is again demonstrated by the
fitted log-normal distributions.
Topic means ranged from 7.9 to 12.9, and topic standard deviations ranged
from 17.5 to 23.2.

\textcolor{red}{
Given that the  scores for document relevance that we have observed are 
distributed as seen previously in other tasks where magnitude estimation has been used, 
we can adopt normalizing and aggregating approaches that have been used before in
the literature.
}

@@

Other normalizations have been proposed, 
for example using the median or the arithmetic mean instead of the
geometric mean (especially when zero is an allowed score, so the
geometric mean cannot be used)~\citep{moskowitz:1977}.
Other approaches include carrying out an additional calibration task
that is aimed at understanding the differences in individual scales;
in our scenario, this might be implemented by exploiting the \nkn and
\hkh documents, or in a more implicit way by relying on the maximal
and minimal scores expressed by each worker.

### 2016-09-14 6_results_RQ2.tex
some small changes, and discovery of Section 7 disaster (which I did not fix)
Attempted to fix the missing answer to RQ2.

In our data, there are
at least ten workers judging the same topic-document pair, and
the source of internal disagreement might be threefold:
(i) the arbitrary, personal scale used by a judge when expressing an ME
score; 
(ii) the subjective nature of relevance; and 
(iii) the often-feared low quality of work in crowdsourcing exercises.
The first has already been discussed and is removed by the geometric
average normalization process (Section~\ref{sec:score-normalization}).
The second is a feature that one might not want to remove from the
data: different judges can truly have different opinions on the
relevance of a document, whatever scale of measurement is used.
Usually IR evaluators assume that this variation is distilled into a single 
value (for example: mean, median or mode) 
that represents a population of users.
Of course, the third source is in need of particular attention: low
quality workers might express unreliable scores, and in general this
would lead to low internal agreement.

@@

While it is difficult to make conclusions about the absolute values of
the ratio or geometric standard deviation without some reference point, it does highlight, for
example, that there are 23 documents that seem to have an
unusually high ratio of 10000 or more.
Looking at each of these cases, there are two distinct causes.
Firstly, 15 of the 23 looked like genuine attempts to assign ME scores,
but the ranges of scores for each worker involved was very wide.
One reason given in the comments by workers was that there was no zero
available, so for documents that were truly off topic, they chose an
arbitrary, very very small, number, making a wide scale.
These wide scales were not markedly compressed by the normalization
stage.
In some cases the numbers assigned seemed arbitrary, but suitably
large or small in correlation with the expert judgments.
The second cause, in 8 of the 23 cases, seemed to be workers not
following instructions or reading documents carefully.
In these cases the comments provided by the workers bore no
relationship to their numeric scores, and the numbers themselves
seemed arbitrary.
We will discuss the seemingly arbitrary nature of some scores later on
(Section~\ref{sec:sig-figs}).

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=.7\linewidth]{figs/pairwise5.pdf}
  \vspace{-0.0cm}
  \caption{Raw agreement on the ordering of relevance of all document pairs (one
small dot per topic) between judges as indicated on the x-axis (ME: magnitude
estimation). The large (red) dot is the mean over all topics in each case. 
    Unjudged documents are excluded.
  \label{fig:agreement}
}
\end{figure}

@@

Figure~\ref{fig:agreement} shows the agreement data stratified by each of the possible six pairs of
ordinal relevance levels of the Sormunen judgments (first six columns)
and ungrouped (seventh column).
The TREC vs.\ ME comparison, aggregated over all topics, is shown in the
eighth column, and for comparison, the final column shows the pairwise
agreement between TREC and Sormunen, assuming that 
\nn equates  to TREC category ``0'', and the other three Sormunen categories 
equate to TREC category ''1''.
Red circles indicate the mean score over the 18 topics for each group.
It can be seen that the rates of agreement are highly consistent when
comparing any of the three relevance scales.
In particular, ME scores for \hh documents are greater than those for 
\nn documents 92\% of the time (mean, first column), which is
higher than the average agreement between TREC and Sormunen
(85\%---mean, last column).
Similarly, the proportion of document pairs that have different TREC
categories are ranked in the same order by ME scores 86\% of the time
(mean, second last column).
%The agreement between 
ME  and Sormunen ranking for documents in the \nn\rr category
also agree 82\% of the time (mean, second column), and over all pairs
Sormunen and ME agree 80\% of the time (mean, seventh column).
This overall average is reduced by the lower agreements between pairs
of documents that are deemed relevant and one or other are in the 
\mm or \rr categories (columns three to six).
This is perhaps unsurprising, as distinguishing ``marginal relevance'' (\mm)
from ``irrelevant'' (\nn) or ``relevant'' (\rr) can be difficult; and likewise 
distinguishing \rr from highly relevant (\hh) can also be challenging.

@@

>>>>

In our data, there are
at least ten workers judging the same topic-document pair, and
the source of internal disagreement might be threefold:
(i) the arbitrary, personal scale used by a judge when expressing an ME
score; 
(ii) the subjective nature of relevance; and 
(iii) the often-feared low quality of work in crowdsourcing exercises.
The first has already been discussed and is removed by the geometric
average normalization process (Section~\ref{sec:score-normalization}).
The second is a feature that one might not want to remove from the
data: different judges can truly have different opinions on the
relevance of a document, whatever scale of measurement is used.
Usually IR evaluators assume that this variation is distilled into a single 
value (for example: mean, median or mode) 
that represents a population of users \sm{, but there might not be a unique truth~\citep{aroyo2015truth}}.
Of course, the third source is in need of particular attention: low
quality workers might express unreliable scores, and in general this
would lead to low internal agreement.

@@

While it is difficult to make conclusions about the absolute values of
the ratio or geometric standard deviation without some reference point, it does highlight, for
example, that there are 23 documents that seem to have an
unusually high ratio of 10000 or more.
Looking at each of these cases, there are two distinct causes.
Firstly, 15 of the 23 looked like genuine attempts to assign ME scores,
but the range of scores for each worker involved was very wide.
One reason given in the comments by workers was that there was no zero
available, so for documents that were truly off topic, they chose an
arbitrary, very very small, number, making a wide scale.
These wide scales were not markedly compressed by the normalization
stage.
In some cases the numbers assigned seemed arbitrary, but suitably
large or small in correlation with the expert judgments.
The second cause, in 8 of the 23 cases, seemed to be workers not
following instructions or reading documents carefully.
In these cases the comments provided by the workers bore no
relationship to their numeric scores, and the numbers themselves
seemed arbitrary.
We will discuss the seemingly arbitrary nature of some scores later on
(Section~\ref{sec:sig-figs}).

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=.75\linewidth]{figs/pairwise5.pdf}
  \vspace{-0.0cm}
  \caption{Raw agreement on the ordering of relevance of all document pairs (one
small dot per topic) between judges as indicated on the x-axis (ME: magnitude
estimation). The large (red) dot is the mean over all topics in each case. 
    Unjudged documents are excluded.
  \label{fig:agreement}
}
\end{figure}

@@

Figure~\ref{fig:agreement} shows the agreement data stratified by each of the possible six pairs of
ordinal relevance levels of the Sormunen judgments (first six columns)
and ungrouped (seventh column).
The TREC vs.\ ME comparison, aggregated over all topics, is shown in the
eighth column, and for comparison, the final column shows the pairwise
agreement between TREC and Sormunen, assuming that 
\nn equates  to TREC category ``0'', and the other three Sormunen categories 
equate to TREC category ''1''.
Red large circles indicate the mean score over the 18 topics for each group.
It can be seen that the rates of agreement are highly consistent when
comparing any of the three relevance scales.
In particular, ME scores for \hh documents are greater than those for 
\nn documents 92\% of the time (mean, first column), which is
higher than the average agreement between TREC and Sormunen
(85\%---mean, last column).
Similarly, the proportion of document pairs that have different TREC
categories are ranked in the same order by ME scores 86\% of the time
(mean, second last column).
%The agreement between 
ME  and Sormunen ranking for documents in the \nn\rr category
also agree 82\% of the time (mean, second column), and over all pairs
Sormunen and ME agree 80\% of the time (mean, seventh column).
This overall average is reduced by the lower agreements between pairs
of documents that are deemed relevant and one or other are in the 
\mm or \rr categories (columns three to six).
This is perhaps unsurprising, as distinguishing ``marginal relevance'' (\mm)
from ``irrelevant'' (\nn) or ``relevant'' (\rr) can be difficult; and likewise 
distinguishing \rr from highly relevant (\hh) can also be challenging.

@@

\subsection{Summary}
\label{sec:summary-1}

\sm{Let us wrap up an overall answer to  RQ\ref{item:rq4}. As we anticipated, using ME with crowdsourcing could potentially raise several concerns, including: subjectiveness of the scale, normalization issues, lack of proper training, potential internal and/or external disagreement, and overall cost in terms of the number of redundant workers needed.}
\sm{On the basis of the results presented so far we can conclude that, at least with our setting and our quality checks in place (see items (a)--(d) in Section~\ref{sec:quality-checks}), we can be reassured that magnitude estimation relevance scores gathered by means of crowdsourcing are reliable. Also, the cost is competitive, as in our experiment it is in the same range of the cost of collecting ordinal scale judgments, and there are clues that it can be lowered by collecting fewer judgments.}
